---
title: "Classification"
author: "Kenneth L Osborne"
date: "August 26, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
suppressPackageStartupMessages(library(magrittr))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(lubridate))
```

# Classification

Classification

You have explored and analyzed customer data collected by the Adventure Works Cycles company. Now you should be ready to apply what you have learned about the data to building, testing, and optimizing a predictive machine learning model.

Specifically, you must use any combination of Azure Machine Learning, R or Python to create a classification model that predicts whether or not a new customer will buy a bike.

Challenge Instructions
To complete this challenge:

1. Use the Adventure Works Cycles customer data you worked with in challenge 1 to create a classification model that predicts whether or not a customer will purchase a bike. The model should predict bike purchasing for new customers for whom no information about average monthly spend or previous bike purchases is available.
2. Download the test data. This data includes customer features but does not include bike purchasing or average monthly spend values.
3. Use your model to predict the corresponding test dataset. Don't forget to apply what you've learned throughout this course.
4. Go to the next page to check how well your prediction against the actual result.

